---
date: 2017-04-25T19:06:58Z
description: "Animating plots with gganimate and tweenr"
draft: true
tags: []
title: Animating plots with gganimate and tweenr
topics: []
header-includes:
   - \usepackage{animate}
---

Animation can be added to ```ggplot()``` plots using the ```gganimate``` package by [David Robinson](https://github.com/dgrtwo/gganimate).

Using the CSO Statbank API __insert link__, we can load in a table of population by county in Ireland for each Census since 1841.

```{r, warning = FALSE, message = FALSE}
library(rjstat)
library(tidyverse)
library(gganimate)
library(tweenr)
library(scales)

census_url <- "http://www.cso.ie/StatbankServices/StatbankServices.svc/jsonservice/responseinstance/EY001"

pop_data <- fromJSONstat(census_url)

pop_data <- pop_data[[1]] %>% 
    as_data_frame()
```

To look at county populations, filter to both sexes and filter out the total state populations.

```{r, warning = FALSE, message = FALSE}
pop_data <- pop_data %>% 
    filter(Sex == "Both sexes", County != "State") %>% 
    rename(year = `Census Year`, pop = value) %>% 
    select(County, year, pop) %>% 
    mutate(year = as.numeric(year))
```

```{r, warning = FALSE, message = FALSE}
pop_edit <- pop_data %>% 
    rename(y = pop, time = year, id = County) %>%
    mutate(ease = "linear")
```

```{r, warning = FALSE, message = FALSE}
pop_tween <- tween_elements(pop_edit, "time", "id", "ease", nframes = 300)

pop_tween <- pop_tween %>%
    mutate(year = round(time), county = .group)

pop_tween <- inner_join(pop_tween, pop_data) %>% 
    as_data_frame()
```

```{r, warning = FALSE, message = FALSE}
county_colours <- read_csv("county-colours.csv")

fill <- setNames(as.character(county_colours$fill), county_colours$county)
line <- setNames(as.character(county_colours$line), county_colours$county)
```

__EDIT:__ It seems that the plot will not render as a gif on this website. I'll need to find a solution to this!

```{r, warning = FALSE, message = FALSE}
p <- pop_tween %>% 
    #filter(year == i) %>%
    transform(county = reorder(county, -pop)) %>%
    ggplot(aes(county, y, colour = county, fill = county, frame = .frame)) +
    geom_bar(stat = "identity", position = "identity", size = 0.2) +
    geom_text(aes(label = year), colour = "black", size = 0.1) +
    scale_fill_manual(name = "county", values = fill) +
    scale_colour_manual(name = "county", values = line) +
    scale_y_continuous(labels = comma) +
    labs(x = "", y = "National Population") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          legend.position="none",
          panel.background = element_blank(),
          panel.grid = element_blank(),
          plot.title = element_text(size = 20))
    
gganimate(p, title_frame = FALSE, saver = "gif", interval = 0.2)
```


